\section{Appendix: Implementational Details}

The model was implemented using the Leabra framework, which is
described in detail in \incite{OReillyMunakata00} and
\incite{OReilly01}, and summarized here.  See
Table~\ref{tab.sim_params} for a listing of parameter values, nearly
all of which are at their default settings.  These same parameters and
equations have been used to simulate over 40 different models in
\incite{OReillyMunakata00}, and a number of other research models.
Thus, the model can be viewed as an instantiation of a systematic
modeling framework using standardized mechanisms, instead of
constructing new mechanisms for each model.  The model can be obtained
by emailing the first author at oreilly@psych.colorado.edu.

\subsection{Point Neuron Activation Function} 

\begin{table}
  \centering
  \begin{tabular}{ll|ll} \hline
Parameter & Value & Parameter & Value \\ \hline
$E_l$ & 0.15 & $\overline{g_l}$ & 0.10 \\
$E_i$ & 0.15 & $\overline{g_i}$ & 1.0 \\
$E_e$ & 1.00 & $\overline{g_e}$ & 1.0 \\
$V_{rest}$ & 0.15 & $\Theta$    & 0.25 \\
$\tau$ & .02 & $\gamma$ & 600 \\
$k$ Post Ctx & 2* & $k$ Output  & 1* \\
$k$ Feat PFC & 2* & $k$ Dim PFC & 1* \\
$k_{hebb}$ & .02 & $\epsilon$ & .01 \\
to AC $\epsilon$ & .04* & \\ \hline
  \end{tabular}
  \caption{\small Parameters for the simulation (see equations in text
    for explanations of parameters).  All are standard
    default parameter values except for those with a * (most of which
    have no default because they are intrinsically task-dependent).
    The faster learning rate ($\epsilon$) for connections into the AC
    was important for ensuring rapid learning of reward.}
  \label{tab.sim_params}
\end{table}

Leabra uses a {\em point neuron} activation function that models the
electrophysiological properties of real neurons, while simplifying
their geometry to a single point.  The membrane potential $V_m$ is
updated as a function of ionic conductances $g$ with reversal
(driving) potentials $E$ as follows:
\begin{equation}
  \Delta V_m(t) = \tau \sum_c g_c(t) \overline{g_c} (E_c - V_m(t))
  \label{eq.vm}
\end{equation}
with 3 channels ($c$) corresponding to: $e$ excitatory input; $l$ leak
current; and $i$ inhibitory input.  Following electrophysiological
convention, the overall conductance is decomposed into a time-varying
component $g_c(t)$ computed as a function of the dynamic state of the
network, and a constant $\overline{g_c}$ that controls the relative
influence of the different conductances.

The excitatory net input/conductance $g_e(t)$ or $\eta_j$ is computed
as the proportion of open excitatory channels as a function of sending
activations times the weight values:
\begin{equation}
  \eta_j = g_e(t) = \langle x_i \wij \rangle = \oneo{n} \sum_i x_i \wij
  \label{eq.net_in_avg}
\end{equation}
The inhibitory conductance is computed via the kWTA function described
in the next section, and leak is a constant.

Activation communicated to other cells ($y_j$) is a thresholded
($\Theta$) sigmoidal function of the membrane potential with gain
parameter $\gamma$:
\begin{equation}
  y_j(t) = \oneo{\left(1 + \oneo{\gamma [V_m(t) - \Theta]_+} \right)}
\end{equation}
where $[x]_+$ is a threshold function that returns 0 if $x<0$ and $x$
if $X>0$.  Note that if it returns 0, we assume $y_j(t) = 0$, to avoid
dividing by 0.  To produce a less discontinuous deterministic
function with a softer threshold, the function is convolved with a
Gaussian noise kernel ($\mu=0$, $\sigma=.005$), which reflects the
intrinsic processing noise of biological neurons:
\begin{equation}
  y^*_j(x) = \int_{-\infty}^{\infty} \oneo{\sqrt{2 \pi} \sigma}
  e^{-z^2/(2 \sigma^2)} y_j(z-x) dz
  \label{eq.conv}
\end{equation}
where $x$ represents the $[V_m(t) - \Theta]_+$ value, and $y^*_j(x)$
is the noise-convolved activation for that value.  In the simulation,
this function is implemented using a numerical lookup table.

\subsection{k-Winners-Take-All Inhibition}

Leabra uses a kWTA (k-Winners-Take-All) function to achieve inhibitory
competition among units within a layer (area).  The kWTA function
computes a uniform level of inhibitory current $g_i$ for all units in
the layer, such that the $k+1$th most excited unit within a layer is
generally below its firing threshold, while the $k$th is typically
above threshold:
\begin{equation}
  g_i = g^{\Theta}_{k+1} + q (g^{\Theta}_k - g^{\Theta}_{k+1})
  \label{eq.g_i}
\end{equation}
where $0<q<1$ (.25 default used here) is a parameter for setting the
inhibition between the upper bound of $g^{\Theta}_k$ and the lower
bound of $g^{\Theta}_{k+1}$.  These boundary inhibition values are
computed as a function of the level of inhibition necessary to keep a
unit right at threshold:
\begin{equation}
  g_i^{\Theta} = \frac{g^*_e \bar{g_e} (E_e - \Theta) +
    g_l \bar{g_l} (E_l - \Theta)}{\Theta - E_i}
  \label{eq.i_at_thr}
\end{equation}
where $g^*_e$ is the excitatory net input without the bias weight
contribution --- this allows the bias weights to override the kWTA
constraint.

In the basic version of the kWTA function, which is relatively rigid
about the kWTA constraint and is therefore used for output layers,
$g^{\Theta}_k$ and $g^{\Theta}_{k+1}$ are set to the threshold
inhibition value for the $k$th and $k+1$th most excited units,
respectively.  In the {\em average-based} kWTA version, $g^{\Theta}_k$
is the average $g_i^{\Theta}$ value for the top $k$ most excited
units, and $g^{\Theta}_{k+1}$ is the average of $g_i^{\Theta}$ for the
remaining $n-k$ units.  This version allows for more flexibility in
the actual number of units active depending on the nature of the
activation distribution in the layer.

\subsection{Hebbian and Error-Driven Learning} 

For learning, Leabra uses a combination of error-driven and Hebbian
learning.  The error-driven component is the symmetric midpoint
version of the GeneRec algorithm \cite{OReilly96}, which is
functionally equivalent to the deterministic Boltzmann machine and
contrastive Hebbian learning (CHL).  The network settles in two
phases, an expectation (minus) phase where the network's actual output
is produced, and an outcome (plus) phase where the target output is
experienced, and then computes a simple difference of a pre and
postsynaptic activation product across these two phases.  For Hebbian
learning, Leabra uses essentially the same learning rule used in
competitive learning or mixtures-of-Gaussians which can be seen as a
variant of the Oja normalization \cite{Oja82}.  The error-driven and
Hebbian learning components are combined additively at each connection
to produce a net weight change.

The equation for the Hebbian weight change is:
\begin{equation}
  \Delta_{hebb} \wij = x^+_i y^+_j - y^+_j \wij = y^+_j (x^+_i - \wij)
  \label{eq.hebb}
\end{equation}
and for error-driven learning using CHL:
\begin{equation}
  \Delta_{err} \wij = (x^+_i y^+_j) - (x^-_i y^-_j)
  \label{eq.err}
\end{equation}
which is subject to a soft-weight bounding to keep within the $0-1$
range:
\begin{equation}
  \Delta_{sberr} \wij = [\Delta_{err}]_+ (1-\wij) + [\Delta_{err}]_- \wij
  \label{eq.err_soft_bound}
\end{equation}
The two terms are then combined additively with a normalized mixing
constant $k_{hebb}$:
\begin{equation}
  \Delta \wij = \epsilon[k_{hebb} (\Delta_{hebb}) + (1-k_{hebb}) (\Delta_{sberr})]
  \label{eq.k_hebb}
\end{equation}

\subsection{Weight Contrast Enhancement} 

One limitation of the Hebbian learning algorithm is that the weights
linearly reflect the strength of the conditional probability.  This
linearity can limit the network's ability to focus on only the
strongest correlations, while ignoring weaker ones.  To remedy this
limitation, we introduce a contrast enhancement function that
magnifies the stronger weights and shrinks the smaller ones in a
parametric, continuous fashion.  This contrast enhancement is achieved
by passing the linear weight values computed by the learning rule
through a sigmoidal nonlinearity of the following form:
\begin{equation}
  \hat{w}_{ij} = \oneo{1 + \left(\frac{\wij}{\theta (1-\wij)}\right)^{-\gamma}}
  \label{eq.wt_off}
\end{equation}
where $\hat{w}_{ij}$ is the contrast-enhanced weight value, and the
sigmoidal function is parameterized by an offset $\theta$ and a gain
$\gamma$ (standard defaults of 1.25 and 6, respectively, used here).  

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
