\documentclass[defaultstyle,12pt]{thesis}
\usepackage{subfiles}
\usepackage{apa} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{url}
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Front matter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{What happens next and when ``next'' happens: \\ Mechanisms of spatial and temporal prediction}
\author{Dean R.}{Wyatte}
\otherdegrees{B.S., Indiana University Bloomington, 2007 \\
M.A., University of Colorado Boulder, 2010}
\degree{Doctor of Philosophy}{Ph.D., Cognitive Neuroscience}
\dept{Department of}{Psychology and Neuroscience}
\advisor{Prof.}{Randall C. O'Reilly}
\reader{Prof.~Tim Curran}	
\readerThree{Prof.~Albert Kim}		%  3rd person to sign thesis
\readerFour{Prof.~Joshua Correll}		%  4th person to sign thesis
\readerFive{Prof.~Nikolaus Correll}		%  5th person to sign thesis
%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abstract{  \OnePageChapter
The physics of the environment provide a rich spatiotemporal structure for our experience. Objects move in predictable ways and their features and identity remain stable across time and space. How does the brain leverage this structure to make predictions about and learn from the environment? This thesis describes research centered around a mechanistic description of sensory prediction called LeabraTI (TI: Temporal Integration) that explains precisely how predictive processing is accomplished in neocortical microcircuits. The fundamental prediction of LeabraTI is that predictions and sensations are interleaved across the same neural tissue at an overall rate of 10 Hz, corresponding to the widely studied alpha rhythm of posterior cortex. Experiments described herein tested this prediction by manipulating the spatiotemporal properties of three-dimensional object stimuli in a laboratory setting. EEG results indicated that predictions were subserved by $\sim$10 Hz oscillations that reliably tracked the onset of stimuli and differentiated between spatially predictable and unpredictable object sequences. There was a behavioral advantage for combined spatial and temporal predictability for discrimination of unlearned objects, but prolonged study of objects under this combined predictability context impaired discriminability relative to other learning contexts. This counterintuitive pattern of results was accounted for by a neural network model that learned three-dimensional viewpoint invariance with LeabraTI's spatiotemporal prediction rule. Synaptic weight scaling from prolonged learning built viewpoint invariance, but led to confusion between ambiguous views of objects, producing slightly lower performance on average. Overall, this work advances a biological architecture for sensory prediction accompanied by empirical evidence that supports learning of realistic time- and space-varying inputs.
}
%%%%%%%%

\acknowledgements{	\OnePageChapter
Many thanks are in order to:
%
\begin{itemize}
\item Randy O'Reilly for his guidance throughout my graduate career. His experience and good intuition (not to mention his unabashed optimism) shaped the way in which I now approach problems as a computational modeler and general scientist. 

\item Tim Curran and Al Kim for their knowledge in areas where Randy lacked expertise -- especially experimental work (Randy proudly claims to have never personally run an experiment).

\item Tim Curran's research assistants. A special thanks to Chris Bird and Krystin Corby for coordinating everything and expediting my experiments in the face of other important deadlines.

\item The many friends I have made through mountain biking (Jim, Heather, Matt, Caroline, Greg, Keith, and others). Thanks for joining me on countless adventures and sorry that they always seemed to end with getting lost deep in the woods.

\item My wife, Debra -- a never ending source of emotional (and financial) support. Thanks for keeping me grounded in reality so that I could see the bigger picture of my work.

\end{itemize}
}
	
\emptyLoT	% use this if there is no List of Tables

\begin{document}

\subfile{chap_intro.tex}
\subfile{chap_leabrati.tex}
\subfile{chap_pleast.tex}
\subfile{chap_bpleast.tex}
\subfile{chap_sims.tex}
\subfile{chap_discuss.tex}

\newpage
\bibliographystyle{apa}
\bibliography{ccnlab}

\end{document}